---
title: "Business Intelligence Lab Submission Markdown"
author: "korn"
date: "8/10/23"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

+---------------------------------------------------+-----------------------------------------------------+
| **Student ID Numbers and Names of Group Members** | 1.  134644 - C - Sebastian Muramara                 |
|                                                   |                                                     |
|                                                   | 2.  136675 - C - Bernard Otieno                     |
|                                                   |                                                     |
|                                                   | 3.  131589 - C - Agnes Anyango                      |
|                                                   |                                                     |
|                                                   | 4.  131582 - C - Njeri Njuguna                      |
|                                                   |                                                     |
|                                                   | 5.  136009 - C- Sera Ndabari                        |
+---------------------------------------------------+-----------------------------------------------------+
| **GitHub Classroom Group Name**                   | Korn                                                |
+---------------------------------------------------+-----------------------------------------------------+
| **Course Code**                                   | BBT4206                                             |
+---------------------------------------------------+-----------------------------------------------------+
| **Course Name**                                   | Business Intelligence II                            |
+---------------------------------------------------+-----------------------------------------------------+
| **Program**                                       | Bachelor of Business Information Technology         |
+---------------------------------------------------+-----------------------------------------------------+
| **Semester Duration**                             | 21^st^ August 2023 to 28^th^ November 2023          |
+---------------------------------------------------+-----------------------------------------------------+

# Setup Chunk

We start by installing all the required packages

```{r Install Packages, echo=TRUE, message=FALSE, warning=FALSE}
## formatR - Required to format R code in the markdown ----
if (!is.element("formatR", installed.packages()[, 1])) {
  install.packages("formatR", dependencies = TRUE,
                   repos="https://cloud.r-project.org")
}
require("formatR")


## readr - Load datasets from CSV files ----
if (!is.element("readr", installed.packages()[, 1])) {
  install.packages("readr", dependencies = TRUE,
                   repos="https://cloud.r-project.org")
}
require("readr")
library(dplyr)
library(magrittr)
library(knitr)
library(ggplot2)
library (tidytext)
library(kableExtra)

```

------------------------------------------------------------------------

**Note:** the following "*KnitR*" options have been set as the defaults in this markdown:\
`knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE, collapse = FALSE, tidy.opts = list(width.cutoff = 80), tidy = TRUE)`.

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
	eval = TRUE,
	echo = TRUE,
	warning = FALSE,
	collapse = FALSE,
	tidy = TRUE
)
```

------------------------------------------------------------------------

**Note:** the following "*R Markdown*" options have been set as the defaults in this markdown:

> output:\
> \
> github_document:\
> toc: yes\
> toc_depth: 4\
> fig_width: 6\
> fig_height: 4\
> df_print: default\
> \
> editor_options:\
> chunk_output_type: console

# Loading the Student Performance Dataset

The 20230412-20230719-BI1-BBIT4-1-StudentPerformanceDataset is then loaded. The dataset and its metadata are available here: <https://drive.google.com/drive/folders/1-BGEhfOwquXF6KKXwcvrx7WuZXuqmW9q?usp=sharing>

```{r Load Dataset, error = TRUE}
student_performance_dataset <-
  readr::read_csv(
                  "../data/20230412-20230719-BI1-BBIT4-1-StudentPerformanceDataset.CSV", # nolint

                  col_types =
                  readr::cols(
                              class_group =
                              readr::col_factor(levels = c("A", "B", "C")),
                              gender = readr::col_factor(levels = c("1", "0")),
                              YOB = readr::col_date(format = "%Y"),
                              regret_choosing_bi =
                              readr::col_factor(levels = c("1", "0")),
                              drop_bi_now =
                              readr::col_factor(levels = c("1", "0")),
                              motivator =
                              readr::col_factor(levels = c("1", "0")),
                              read_content_before_lecture =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              anticipate_test_questions =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              answer_rhetorical_questions =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              find_terms_I_do_not_know =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              copy_new_terms_in_reading_notebook =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              take_quizzes_and_use_results =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              reorganise_course_outline =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              write_down_important_points =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              space_out_revision =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              studying_in_study_group =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              schedule_appointments =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              goal_oriented =
                              readr::col_factor(levels =
                                                c("1", "0")),
                              spaced_repetition =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4")),
                              testing_and_active_recall =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4")),
                              interleaving =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4")),
                              categorizing =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4")),
                              retrospective_timetable =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4")),
                              cornell_notes =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4")),
                              sq3r = readr::col_factor(levels =
                                                       c("1", "2", "3", "4")),
                              commute = readr::col_factor(levels =
                                                          c("1", "2",
                                                            "3", "4")),
                              study_time = readr::col_factor(levels =
                                                             c("1", "2",
                                                               "3", "4")),
                              repeats_since_Y1 = readr::col_integer(),
                              paid_tuition = readr::col_factor(levels =
                                                               c("0", "1")),
                              free_tuition = readr::col_factor(levels =
                                                               c("0", "1")),
                              extra_curricular = readr::col_factor(levels =
                                                                   c("0", "1")),
                              sports_extra_curricular =
                              readr::col_factor(levels = c("0", "1")),
                              exercise_per_week = readr::col_factor(levels =
                                                                    c("0", "1",
                                                                      "2",
                                                                      "3")),
                              meditate = readr::col_factor(levels =
                                                           c("0", "1",
                                                             "2", "3")),
                              pray = readr::col_factor(levels =
                                                       c("0", "1",
                                                         "2", "3")),
                              internet = readr::col_factor(levels =
                                                           c("0", "1")),
                              laptop = readr::col_factor(levels = c("0", "1")),
                              family_relationships =
                              readr::col_factor(levels =
                                                c("1", "2", "3", "4", "5")),
                              friendships = readr::col_factor(levels =
                                                              c("1", "2", "3",
                                                                "4", "5")),
                              romantic_relationships =
                              readr::col_factor(levels =
                                                c("0", "1", "2", "3", "4")),
                              spiritual_wellnes =
                              readr::col_factor(levels = c("1", "2", "3",
                                                           "4", "5")),
                              financial_wellness =
                              readr::col_factor(levels = c("1", "2", "3",
                                                           "4", "5")),
                              health = readr::col_factor(levels = c("1", "2",
                                                                    "3", "4",
                                                                    "5")),
                              day_out = readr::col_factor(levels = c("0", "1",
                                                                     "2", "3")),
                              night_out = readr::col_factor(levels = c("0",
                                                                       "1", "2",
                                                                       "3")),
                              alcohol_or_narcotics =
                              readr::col_factor(levels = c("0", "1", "2", "3")),
                              mentor = readr::col_factor(levels = c("0", "1")),
                              mentor_meetings = readr::col_factor(levels =
                                                                  c("0", "1",
                                                                    "2", "3")),
                              `Attendance Waiver Granted: 1 = Yes, 0 = No` =
                              readr::col_factor(levels = c("0", "1")),
                              GRADE = readr::col_factor(levels =
                                                        c("A", "B", "C", "D",
                                                          "E"))),
                  locale = readr::locale())

View(student_performance_dataset)

```

## Description of the Dataset

We then display the number of observations and number of variables. We have 101 observations and 100 variables to work with.

```{r Your Fourth Code Chunk}
dim(student_performance_dataset)
```

Next, we display the quartiles for each numeric variable[*... think of this process as **"storytelling using the data."** Tell us what is happening; tell us what you are discovering as you proceed with the markdown; walk us through your code step-by-step (a code walkthrough).*]{#highlight style="color: blue"}

```{r Your Fifth Code Chunk}
summary(student_performance_dataset)
```

# Average rating per gender and per group

Here we obtain data of the average rating per gender and per group and create a bar chart visualizing the results from Group A to Group C.

```{r Your Sixth Code Chunk}
# Fill this with R related code that will be executed when the R markdown file

#blue_grey_theme
blue_grey_colours_11 <- c("#27408E", "#304FAF", "#536CB5", "#6981c7", "#8da0db",
                          "#dde5ec", "#c8c9ca", "#B9BCC2", "#A7AAAF", "#888A8E",
                          "#636569")

blue_grey_colours_6 <- c("#27408E", "#304FAF", "#536CB5",
                         "#B9BCC2", "#A7AAAF", "#888A8E")

blue_grey_colours_4 <- c("#27408E", "#536CB5",
                         "#B9BCC2", "#888A8E")

blue_grey_colours_3 <- c("#6981c7", "#304FAF", "#888A8E")

blue_grey_colours_2 <- c("#27408E",
                         "#888A8E")

blue_grey_colours_1 <- c("#6981c7")

# Custom theme for visualizations
blue_grey_theme <- function() {
  theme(
    axis.ticks = element_line(
                              linewidth = 1, linetype = "dashed",
                              lineend = NULL, color = "#dfdede",
                              arrow = NULL, inherit.blank = FALSE),
    axis.text = element_text(
                             face = "bold", color = "#3f3f41",
                             size = 12, hjust = 0.5),
    axis.title = element_text(face = "bold", color = "#3f3f41",
                              size = 14, hjust = 0.5),
    plot.title = element_text(face = "bold", color = "#3f3f41",
                              size = 16, hjust = 0.5),
    panel.grid = element_line(
                              linewidth = 0.1, linetype = "dashed",
                              lineend = NULL, color = "#dfdede",
                              arrow = NULL, inherit.blank = FALSE),
    panel.background = element_rect(fill = "#f3eeee"),
    legend.title = element_text(face = "plain", color = "#3f3f41",
                                size = 12, hjust = 0),
    legend.position = "right"
  )
}


# evaluation_per_group_per_gender
evaluation_per_group_per_gender <- student_performance_dataset %>% # nolint
  mutate(`Student's Gender` =
           ifelse(gender == 1, "Male", "Female")) %>%
  select(class_group, gender,
         `Student's Gender`, `Average Course Evaluation Rating`) %>%
  filter(!is.na(`Average Course Evaluation Rating`)) %>%
  group_by(class_group, `Student's Gender`) %>%
  summarise(average_evaluation_rating =
              mean(`Average Course Evaluation Rating`)) %>%
  arrange(desc(average_evaluation_rating), .by_group = TRUE)




evaluation_per_group_per_gender %>%
  ggplot() +
  geom_bar(aes(x = class_group, y = average_evaluation_rating,
               fill = `Student's Gender`),
           stat = "identity", position = "dodge") +
  expand_limits(y = 0) +
  blue_grey_theme() +
  scale_fill_manual(values = blue_grey_colours_2) +
  ggtitle("Course Evaluation Rating per Group and per Gender") +
  labs(x = "Class Group", y = "Average Rating")
```
## Word Count.

Number of Significant Words in Evaluation Wishes per Group: Minus contractions, special characters, stop-words, short words, and censored words.

```{r Your Seventh Code Chunk}
# Fill this with other R related code that will be executed when the R markdown

#undesirable_words
undesirable_words <- c("wow", "lol", "none", "na")
#evaluation_likes_and_wishes
evaluation_likes_and_wishes <- student_performance_dataset %>%
  mutate(`Student's Gender` =
           ifelse(gender == 1, "Male", "Female")) %>%
  rename(`Class Group` = class_group) %>%
  rename(Likes = `D - 1. \nWrite two things you like about the teaching and learning in this unit so far.`) %>% # nolint
  rename(Wishes = `D - 2. Write at least one recommendation to improve the teaching and learning in this unit (for the remaining weeks in the semester)`) %>% # nolint
  select(`Class Group`,
         `Student's Gender`, `Average Course Evaluation Rating`,
         Likes, Wishes) %>%
  filter(!is.na(`Average Course Evaluation Rating`)) %>%
  arrange(`Class Group`)

#unnest
evaluation_likes_filtered <- evaluation_likes_and_wishes %>% # nolint
  # We start by tokenization (un-nesting words). This is from the variable
  # "Like" into the variable "word".
  unnest_tokens(word, Likes) %>%
  # Then we remove stopwords using an anti-join (remember this from the
  # BBT3104: Advanced Database Systems course)
  # Anti-join: do not join where the word is in the list of stopwords
  anti_join(stop_words, by = c("word")) %>%
  distinct() %>%
  # Censor or filter out unwanted words
  filter(!word %in% undesirable_words) %>%
  # Include only words that are more than 3 characters long (assuming that
  # these are the words that are meaningful)
  filter(nchar(word) > 3) %>%
  # We then rename the variable "word" for ease of use.
  rename(`Likes (tokenized)` = word) %>%
  # We focus only on the likes in this data frame
  select(-Wishes)

#evaluation_wishes_filtered
evaluation_wishes_filtered <- evaluation_likes_and_wishes %>% # nolint
  unnest_tokens(word, Wishes) %>%
  anti_join(stop_words, by = c("word")) %>%
  distinct() %>%
  filter(!word %in% undesirable_words) %>%
  filter(nchar(word) > 3) %>%
  rename(`Wishes (tokenized)` = word) %>%
  select(-Likes)


word_count_per_group_wishes <- evaluation_wishes_filtered %>%
  group_by(`Class Group`) %>%
  summarise(num_words = n()) %>%
  arrange(desc(num_words))

word_count_per_group_wishes %>%
  mutate(num_words = color_bar("lightblue")(num_words)) %>%
  rename(`Number of Words` = num_words) %>%
  kable("html", escape = FALSE, align = "c",
        caption = "Number of Significant Words in Evaluation Wishes 
                   per Group: Minus contractions, special characters, 
                   stopwords, short words, and censored words.") %>%
  kable_styling(bootstrap_options =
                  c("striped", "condensed", "bordered"),
                full_width = FALSE)

```

## Top 10 words per gender

This code generates a horizontal bar chart that displays the top 10 most frequently used words in course evaluation "Likes" for each gender group. The bars are coloured by gender, and each gender has its own facet (sub-plot) for comparison. This visualization helps you understand which words are most commonly used in course evaluation likes and whether there are any gender-specific patterns in the choice of words.

```{r Your Eighth Code Chunk}
# Fill this with other R related code that will be executed when the R markdown

popular_words <- evaluation_likes_filtered %>%
  group_by(`Student's Gender`) %>%
  count(`Likes (tokenized)`, `Student's Gender`, sort = TRUE) %>%
  slice(seq_len(10)) %>%
  ungroup() %>%
  arrange(`Student's Gender`, n) %>%
  mutate(row = row_number())

popular_words %>%
  ggplot(aes(row, n, fill = `Student's Gender`)) +
  geom_col(fill = blue_grey_colours_1) +
  blue_grey_theme() +
  labs(x = "Word in Course Evaluation",
       y = "Number of Times Used (Term Frequency)") +
  ggtitle("Most Frequently Used Words in Course Evaluation Likes per Gender") +
  facet_wrap(~`Student's Gender`, scales = "free") +
  scale_x_continuous(
                     breaks = popular_words$row,
                     labels = popular_words$`Likes (tokenized)`) +
  coord_flip()

```

## Top 10 words per gender

This code generates a horizontal bar chart that displays the top 10 most frequently used words in course evaluation "Wishes" for each gender group. Similar to the one above

```{r}

popular_words <- evaluation_wishes_filtered %>%
  group_by(`Student's Gender`) %>%
  count(`Wishes (tokenized)`, `Student's Gender`, sort = TRUE) %>%
  slice(seq_len(10)) %>%
  ungroup() %>%
  arrange(`Student's Gender`, n) %>%
  mutate(row = row_number())

popular_words %>%
  ggplot(aes(row, n, fill = `Student's Gender`)) +
  geom_col(fill = blue_grey_colours_1) +
  blue_grey_theme() +
  labs(x = "Word in Course Evaluation", y = "Number of Times Used") +
  ggtitle("Most Frequently Used Words in Course Evaluation Wishes per Gender") +
  facet_wrap(~`Student's Gender`, scales = "free") +
  scale_x_continuous(
                     breaks = popular_words$row,
                     labels = popular_words$`Wishes (tokenized)`) +
  coord_flip()
```

## MOST IMPORTANT WORDS BY TF-IDF SCORE IN COURSE EVALUATION CLASS GROUP (LIKES)

These visualizations help identify the most important words in course evaluation "Likes" based on TF-IDF scores, allowing you to understand which words are particularly significant for each gender group. This can provide insights into the specific language and preferences used by different genders in their evaluations

```{r}
popular_tfidf_words_gender_likes <- evaluation_likes_filtered %>% # nolint
  unnest_tokens(word, `Likes (tokenized)`) %>%
  distinct() %>%
  filter(!word %in% undesirable_words) %>%
  filter(nchar(word) > 3) %>%
  rename(`Likes (tokenized)` = word) %>%
  select(`Class Group`, `Student's Gender`,
         `Average Course Evaluation Rating`, `Likes (tokenized)`) %>%
  count(`Student's Gender`, `Likes (tokenized)`, sort = TRUE) %>%
  ungroup() %>%
  bind_tf_idf(`Likes (tokenized)`, `Student's Gender`, n)

head(popular_tfidf_words_gender_likes)

top_popular_tfidf_words <- popular_tfidf_words_gender_likes %>%
  arrange(desc(tf_idf)) %>%
  mutate(`Likes (tokenized)` =
           factor(`Likes (tokenized)`,
                  levels = rev(unique(`Likes (tokenized)`)))) %>%
  group_by(`Student's Gender`) %>%
  slice(seq_len(10)) %>%
  ungroup() %>%
  arrange(`Student's Gender`, tf_idf) %>%
  mutate(row = row_number())

top_popular_tfidf_words %>%
  ggplot(aes(x = row, tf_idf, fill = `Student's Gender`)) +
  geom_col(fill = blue_grey_colours_1) +
  blue_grey_theme() +
  labs(x = "Word in Course Evaluation", y = "TF-IDF Score") +
  ggtitle("Important Words using TF-IDF by Chart Level") +
  ggtitle("Most Important Words by TF-IDF Score in Course Evaluation Likes per 
      Class Group") +
  facet_wrap(~`Student's Gender`, scales = "free") +
  scale_x_continuous(
                     breaks = top_popular_tfidf_words$row,
                     labels = top_popular_tfidf_words$`Likes (tokenized)`) +
  coord_flip()

### TF-IDF Score per Group ----
popular_tfidf_words_likes <- evaluation_likes_filtered %>% # nolint
  unnest_tokens(word, `Likes (tokenized)`) %>%
  distinct() %>%
  filter(!word %in% undesirable_words) %>%
  filter(nchar(word) > 3) %>%
  rename(`Likes (tokenized)` = word) %>%
  select(`Class Group`, `Student's Gender`,
         `Average Course Evaluation Rating`, `Likes (tokenized)`) %>%
  count(`Class Group`, `Likes (tokenized)`, sort = TRUE) %>%
  ungroup() %>%
  bind_tf_idf(`Likes (tokenized)`, `Class Group`, n)

head(popular_tfidf_words_likes)

top_popular_tfidf_words <- popular_tfidf_words_likes %>%
  arrange(desc(tf_idf)) %>%
  mutate(`Likes (tokenized)` =
           factor(`Likes (tokenized)`,
                  levels = rev(unique(`Likes (tokenized)`)))) %>%
  group_by(`Class Group`) %>%
  slice(seq_len(10)) %>%
  ungroup() %>%
  arrange(`Class Group`, tf_idf) %>%
  mutate(row = row_number())

top_popular_tfidf_words %>%
  ggplot(aes(x = row, tf_idf, fill = `Class Group`)) +
  geom_col(fill = blue_grey_colours_1) +
  blue_grey_theme() +
  labs(x = "Word in Course Evaluation", y = "TF-IDF Score") +
  ggtitle("Important Words using TF-IDF by Chart Level") +
  ggtitle("Most Important Words by TF-IDF Score in Course Evaluation Likes per 
      Class Group") +
  facet_wrap(~`Class Group`, scales = "free") +
  scale_x_continuous(
                     breaks = top_popular_tfidf_words$row,
                     labels = top_popular_tfidf_words$`Likes (tokenized)`) +
  coord_flip()


```

## MOST IMPORTANT WORDS BY TF-IDF SCORE IN COURSE EVALUATION CLASS GROUP (WISHES)

code generates visualizations related to the TF-IDF (Term Frequency-Inverse Document Frequency) analysis for course evaluation "Wishes" data, grouped by class group.

```{r}
popular_tfidf_words_likes <- evaluation_wishes_filtered %>% # nolint
  unnest_tokens(word, `Wishes (tokenized)`) %>%
  distinct() %>%
  filter(!word %in% undesirable_words) %>%
  filter(nchar(word) > 3) %>%
  rename(`Wishes (tokenized)` = word) %>%
  select(`Class Group`, `Student's Gender`,
         `Average Course Evaluation Rating`, `Wishes (tokenized)`) %>%
  count(`Class Group`, `Wishes (tokenized)`, sort = TRUE) %>%
  ungroup() %>%
  bind_tf_idf(`Wishes (tokenized)`, `Class Group`, n)

head(popular_tfidf_words_likes)

top_popular_tfidf_words <- popular_tfidf_words_likes %>%
  arrange(desc(tf_idf)) %>%
  mutate(`Wishes (tokenized)` =
           factor(`Wishes (tokenized)`,
                  levels = rev(unique(`Wishes (tokenized)`)))) %>%
  group_by(`Class Group`) %>%
  slice(seq_len(10)) %>%
  ungroup() %>%
  arrange(`Class Group`, tf_idf) %>%
  mutate(row = row_number())

top_popular_tfidf_words %>%
  ggplot(aes(x = row, tf_idf, fill = `Class Group`)) +
  geom_col(fill = blue_grey_colours_1) +
  blue_grey_theme() +
  labs(x = "Word in Course Evaluation", y = "TF-IDF Score") +
  ggtitle("Important Words using TF-IDF by Chart Level") +
  ggtitle("Most Important Words by TF-IDF Score in Course Evaluation Wishes per 
      Class Group") +
  facet_wrap(~`Class Group`, scales = "free") +
  scale_x_continuous(
                     breaks = top_popular_tfidf_words$row,
                     labels = top_popular_tfidf_words$`Wishes (tokenized)`) +
  coord_flip()
```

**etc.** as per the lab submission requirements. Be neat and communicate in a clear and logical manner.
